---
title: "Report: IPC scorecard analysis"
author: "David Niyukuri et Landry Kabego"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)

```




```{r load_packages}

suppressMessages(library(readxl))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(ggforce))
suppressMessages(library(kableExtra))
suppressMessages(library(lubridate))
suppressMessages(library(ggpubr))
suppressMessages(library(linelist))


```


# Data loading and cleaning



```{r read_scripts}

# read scripts
path_to_scripts <- here::here("Utility_Functions")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names=TRUE)
for (file in scripts_files) source(file)

```

<!---
# Chargement des données et nettoyage

--->

```{r load_data}


# path to raw data

path_to_data <- here::here("data", "scorecard-2019-12-13-17-30-48.xlsx") 


# get data file

data_kobo <- rio::import(path_to_data,
                         guess_max = 1e4) %>%
  clean_data(guess_dates = FALSE) %>%
  as_tibble()


# Note: we have two forms to colelct the IPC data: baseline and scorecard



# data with IPC scorecard form
data_scorecard <- dplyr::filter(data_kobo, data_kobo$veuillez_selectionner_le_questionnaire=="ipc_scorecard")

# data with baseline form
data_baseline <- dplyr::filter(data_kobo, data_kobo$veuillez_selectionner_le_questionnaire=="baseline")


# Data for IPC score analysis: the combination of the data from the two forms
data_score <- rbind(data_scorecard, data_baseline)



# scorecard_evaluation_1st_cleaning script

data_scorecard_clean <- scorecard_evaluation_1st_cleaning(data_score = data_score)


```



# Descriptive analysis of recent IPC evaluation


## Overall IPC score of recent evaluation in Health Zones

```{r descr}
# Consider recent evaluation ONLY


data_nziza_descriptive_recent <- recent_evaluation_only_fosa(data_hz = data_scorecard_clean)



# I. DESCRIPTIVE: mean, median, sd - overall, or BY categorie, partners, zone de sante, type structure de sante -------------------



# I.1. Overall by Health Zones

stats_overall <- ggpubr::desc_statby(data_nziza_descriptive_recent, measure.var = "score_pci_100",
                                     grps = "zone_de_sante")
stats_overall_c <- stats_overall[, c("zone_de_sante", "min", "max", "median", "mean", "iqr")]

stats_overall_c[,2:6] <- round(stats_overall_c[,2:6])

# Nbr of health Facilities
zone_de_sante_n <- data_nziza_descriptive_recent %>% group_by(zone_de_sante) %>% tally()

stats_overall_c <- cbind(stats_overall_c, zone_de_sante_n[,2])

names(stats_overall_c) <- c("zone_de_sante", "min", "max", "median", "mean", "iqr", "nbr_healf_fac")

stats_overall_c %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  kableExtra::add_header_above( c( "Table of summary statistics of overall IPC score of recent evaluation in Health Zones" = 7) ) %>% 
  kableExtra::row_spec(17)



# Plot

plot.stats_overall <- ggplot(stats_overall, aes(x=zone_de_sante, y=median, group = zone_de_sante)) + 
  geom_errorbar(aes(ymin=min, ymax=max), width=.1) +
  geom_line(size=.3) +
  geom_point(size=2) + 
  rotate_x_text(45) +
  xlab("Health Zone") + ylab("IPC score")+
  ggtitle("Overall IPC score of recent evaluation in Health Zones")


plot.stats_overall



```

## Overall IPC score of recent evaluation by type of Health Facilities



```{r receHF}

# 1.2. Overall by type of Health Facilities

stats_overall_type_hf <- ggpubr::desc_statby(data_nziza_descriptive_recent, measure.var = "score_pci_100",
                                             grps = "type_structure_de_sante")
stats_overall_type_hf_c <- stats_overall_type_hf[, c("type_structure_de_sante", "min", "max", "median", "mean", "iqr")]

stats_overall_type_hf_c[,2:6] <- round(stats_overall_type_hf_c[,2:6])


# Nbr of Health Facilities
type_structure_de_sante_n <- data_nziza_descriptive_recent %>% group_by(type_structure_de_sante) %>% tally()


stats_overall_type_hf_c <- cbind(stats_overall_type_hf_c, type_structure_de_sante_n[,2])

names(stats_overall_type_hf_c) <- c("zone_de_sante", "min", "max", "median", "mean", "iqr", "nbr_healf_fac")

stats_overall_type_hf_c %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  kableExtra::add_header_above( c( "Table of summary statistics of overall IPC score of recent evaluation by type of Health Facilities" = 7) ) %>% 
  kableExtra::row_spec(9)



# Plot

plot.stats_overall_type_hf <- ggplot(stats_overall_type_hf, aes(x=type_structure_de_sante, y=median, group = type_structure_de_sante)) + 
  geom_errorbar(aes(ymin=min, ymax=max), width=.1) +
  geom_line(size=.3) +
  geom_point(size=2) + 
  rotate_x_text(45) +
  xlab("Types of Health Facilities") + ylab("IPC score")+
  ggtitle("Overall IPC score of recent evaluation by type of Health Facilities")

plot.stats_overall_type_hf


```


## Overall IPC score of recent evaluation by category of Health Facilities



```{r recetn_cat}

# 1.3. Overall by categories of Health Facilities

stats_overall_categorie <- ggpubr::desc_statby(data_nziza_descriptive_recent, measure.var = "score_pci_100",
                                               grps = "categorie")

stats_overall_categorie_c <- stats_overall_categorie[, c("categorie", "min", "max", "median", "mean", "iqr")]

stats_overall_categorie_c[,2:6] <- round(stats_overall_categorie_c[,2:6])

# Nbr of Health Facilities

categorie_n <- data_nziza_descriptive_recent %>% group_by(categorie) %>% tally()


stats_overall_categorie_c <- cbind(stats_overall_categorie_c, categorie_n[,2])


names(stats_overall_categorie_c) <- c("zone_de_sante", "min", "max", "median", "mean", "iqr", "nbr_healf_fac")



stats_overall_categorie_c %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  kableExtra::add_header_above( c( "Table of summary statistics of overall IPC score of recent evaluation by categogry of Health Facilities" = 7) ) %>% 
  kableExtra::row_spec(4)



# Plot

plot.stats_overall_categorie <- ggplot(stats_overall_categorie, aes(x=categorie, y=median,  group = categorie)) + 
  geom_errorbar(aes(ymin=min, ymax=max), width=.1) +
  geom_line(size=.3) +
  geom_point(size=2) + 
  rotate_x_text(45) +
  xlab("Category of Health Facility") + ylab("IPC score")+
  ggtitle("Overall IPC score of recent evaluation by categogry of Health Facilities")


plot.stats_overall_categorie


```



## Overall IPC score of recent evaluation by presence of partners


```{r recent_partn}


# 1.4. Overall by presence of partners

stats_overall_partners <- ggpubr::desc_statby(data_nziza_descriptive_recent, measure.var = "score_pci_100",
                                              grps = "appuyee_par_partenaire")

stats_overall_partners_c <- stats_overall_partners[, c("appuyee_par_partenaire", "min", "max", "median", "mean", "iqr")]

stats_overall_partners_c[,2:6] <- round(stats_overall_partners_c[,2:6])


# Nbr of Health Facilities
partner_status_n <- data_nziza_descriptive_recent %>% group_by(appuyee_par_partenaire) %>% tally()


stats_overall_partners_c <- cbind(stats_overall_partners_c, partner_status_n[,2])


names(stats_overall_partners_c) <- c("zone_de_sante", "min", "max", "median", "mean", "iqr", "nbr_healf_fac")


stats_overall_partners_c %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  kableExtra::add_header_above( c( "Table of summary statistics of overall IPC score of recent evaluation by presence of partners" = 7) ) %>% 
  kableExtra::row_spec(2)


# PLot

plot.stats_overall_partners <- ggplot(stats_overall_partners, aes(x=appuyee_par_partenaire, y=median, colour=appuyee_par_partenaire, group = appuyee_par_partenaire)) + 
  geom_errorbar(aes(ymin=min, ymax=max), width=.1) +
  geom_line(size=.3) +
  geom_point(size=2) + 
  rotate_x_text(45) +
  xlab("Supported by Partners") + ylab("IPC score")+
  ggtitle("Overall IPC score of recent evaluation by presence of partners")

plot.stats_overall_partners


```



## Overall IPC score of recent evaluation by Health Zones and type of Health Facilities


```{r recent_hz_type}

# 1.5. Overall by Health Zones and type of Health Facilities

stats_overall_hz_hf <- ggpubr::desc_statby(data_nziza_descriptive_recent, measure.var = "score_pci_100",
                                           grps = c("zone_de_sante", "type_structure_de_sante"))
stats_overall_hz_hf <- stats_overall_hz_hf[, c("zone_de_sante", "type_structure_de_sante", "min", "max", "median", "mean", "iqr")]

stats_overall_hz_hf[,3:7] <- round(stats_overall_hz_hf[,3:7])


#
#

zone_de_sante_type_structure_de_sante_n <- data_nziza_descriptive_recent %>% group_by(zone_de_sante) %>% count(type_structure_de_sante) 

stats_overall_hz_hf_c <- cbind(stats_overall_hz_hf, zone_de_sante_type_structure_de_sante_n[,3])

names(stats_overall_hz_hf_c) <- c("zone_de_sante", "type_structure_de_sante", "min", "max", "median", "mean", "iqr", "nbr_health_fac")
#
#


stats_overall_hz_hf_c %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  kableExtra::add_header_above( c( "Table of summary statistics of overall IPC score of recent evaluation by Health Zones and type of Health Facilities" = 8) ) %>% 
  kableExtra::row_spec(90)



# Plot

plot.stats_overall_hz_hf_c <- ggplot(stats_overall_hz_hf_c, aes(x=type_structure_de_sante, y=median, colour=type_structure_de_sante, group = type_structure_de_sante)) + 
  geom_errorbar(aes(ymin=min, ymax=max), width=.1) +
  geom_line(size=.3) +
  facet_wrap(.~zone_de_sante) +
  geom_point(size=2) +  
  rotate_x_text(45) +
  xlab("Type of Health Facility") + ylab("IPC score")+
  ggtitle("Overall IPC score of recent evaluation by Health Zones and type of Health Facilities")

plot.stats_overall_hz_hf_c



```



## Overall IPC score of recent evaluation by Health Zones and category of Health Facilities


```{r recent_hz_cat}

# 1.6. Overall by Health Zones and by categories of Health Facilities

stats_overall_hz_cat <- ggpubr::desc_statby(data_nziza_descriptive_recent, measure.var = "score_pci_100",
                                            grps = c("zone_de_sante", "categorie"))
stats_overall_hz_cat_c <- stats_overall_hz_cat[, c("zone_de_sante", "categorie", "min", "max", "median", "mean", "iqr")]

stats_overall_hz_cat_c[,3:7] <- round(stats_overall_hz_cat_c[,3:7])

#
#
zone_de_sante_categorie_n <- data_nziza_descriptive_recent %>% group_by(zone_de_sante) %>% count(categorie) 

stats_overall_hz_cat_c <- cbind(stats_overall_hz_cat_c, zone_de_sante_categorie_n[,3])

names(stats_overall_hz_cat_c) <- c("zone_de_sante", "categorie", "min", "max", "median", "mean", "iqr", "nbr_health_fac")

#
#


stats_overall_hz_cat_c %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  kableExtra::add_header_above( c( "Table of summary statistics of overall IPC score of recent evaluation by Health Zones and by categories of Health Facilities" = 8) ) %>% 
  kableExtra::row_spec(55)



# Plot

plot.stats_overall_hz_cat_c <- ggplot(stats_overall_hz_cat_c, aes(x=categorie, y=median, colour=categorie, group = categorie)) + 
  geom_errorbar(aes(ymin=min, ymax=max), width=.1) +
  geom_line(size=.3) +
  facet_wrap(.~zone_de_sante) +
  geom_point(size=2) +  
  rotate_x_text(45) +
  xlab("Category of Health Facility") + ylab("IPC score")+
  ggtitle("Overall IPC score of recent evaluation by Health Zones and by categories of Health Facilities")


plot.stats_overall_hz_cat_c


```



## Overall IPC score of recent evaluation by Health Zones and by presence of partners



```{r recent_hz_partn}

# 1.7. Overall by Health Zones and by presence of partners

stats_overall_hz_partn <- ggpubr::desc_statby(data_nziza_descriptive_recent, measure.var = "score_pci_100",
                                              grps = c("zone_de_sante", "appuyee_par_partenaire"))

stats_overall_hz_partn_c <- stats_overall_hz_partn[, c("zone_de_sante", "appuyee_par_partenaire", "min", "max", "median", "mean", "iqr")]

stats_overall_hz_partn_c[,3:7] <- round(stats_overall_hz_partn_c[,3:7])

#
#
zone_de_sante_partner_status_n <- data_nziza_descriptive_recent %>% group_by(zone_de_sante) %>% count(appuyee_par_partenaire) 

stats_overall_hz_partn_c <- cbind(stats_overall_hz_partn_c, zone_de_sante_partner_status_n[,3])
#

names(stats_overall_hz_partn_c) <- c("zone_de_sante", "appuyee_par_partenaire", "min", "max", "median", "mean", "iqr", "nbr_health_fac")

#

stats_overall_hz_partn_c %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  kableExtra::add_header_above( c( "Table of summary statistics of overall IPC score of recent evaluation by Health Zones and by presence of partners" = 8) ) %>% 
  kableExtra::row_spec(29)



# Plot

plot.stats_overall_hz_partn_c <- ggplot(stats_overall_hz_partn_c, aes(x=appuyee_par_partenaire, y=median, colour=appuyee_par_partenaire, group = appuyee_par_partenaire)) + 
  geom_errorbar(aes(ymin=min, ymax=max), width=.1) +
  geom_line(size=.3) +
  facet_wrap(.~zone_de_sante) +
  geom_point(size=2) +  
  rotate_x_text(45) +
  xlab("Supported by Partners") + ylab("IPC score")+
  ggtitle("Overall IPC score of recent evaluation by Health Zones and by presence of partners")


plot.stats_overall_hz_partn_c


```



# Descriptive analysis for components of IPC in recent evaluation


Note that the following results are from recent evaluation.


## IPC Component 1 by Health Zone

```{r ipc-cps}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_1 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_1)

plot_c_1 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_1)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 1 per Health Zone")

plot_c_1

```


## IPC Component 2 by Health Zone

```{r ipc-cps_c_2}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_2 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_2)

plot_c_2 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_2)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 2 per Health Zone")

plot_c_2

```



## IPC Component 3 by Health Zone

```{r ipc-cps_c_3}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_3 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_3)

plot_c_3 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_3)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 3 per Health Zone")

plot_c_3

```


## IPC Component 4 by Health Zone

```{r ipc-cps_c_4}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_4 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_4)

plot_c_4 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_4)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 4 per Health Zone")

plot_c_4

```



## IPC Component 5 by Health Zone

```{r ipc-cps_c_5}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_5 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_5)

plot_c_5 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_5)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 5 per Health Zone")

plot_c_5

```



## IPC Component 6 by Health Zone

```{r ipc-cps_c_6}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_6 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_6)

plot_c_6 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_6)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 6 per Health Zone")

plot_c_6

```


## IPC Component 7 by Health Zone

```{r ipc-cps_c_7}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_7 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_7)

plot_c_7 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_7)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 7 per Health Zone")

plot_c_7

```



## IPC Component 8 by Health Zone

```{r ipc-cps_c_8}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_8 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_8)

plot_c_8 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_8)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 8 per Health Zone")

plot_c_8

```



## IPC Component 9 by Health Zone

```{r ipc-cps_c_9}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_9 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_9)

plot_c_9 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_9)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 9 per Health Zone")

plot_c_9

```



## IPC Component 10 by Health Zone

```{r ipc-cps_c_10}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_10 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_10)

plot_c_10 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_10)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 10 per Health Zone")

plot_c_10

```



## IPC Component 11 by Health Zone

```{r ipc-cps_c_11}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_11 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_11)

plot_c_11 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_11)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 11 per Health Zone")

plot_c_11

```



## IPC Component 12 by Health Zone

```{r ipc-cps_c_12}

data_nziza_descriptive_recent_ipc_cs <- data_nziza_descriptive_recent

data_nziza_descriptive_recent_ipc_cs$score_c_12 <- factor(data_nziza_descriptive_recent_ipc_cs$score_c_12)

plot_c_12 <- ggplot(data_nziza_descriptive_recent_ipc_cs, aes(x = zone_de_sante, fill = score_c_12)) +
  geom_bar() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Score level", breaks = c("0", "1", "2", "3"), labels = c("0", "1", "2", "3")) + 
  xlab("Health Zone") +
  ylab("Count") +  
  ggtitle("Score for IPC component 12 per Health Zone")

plot_c_12

```



## IPC scores for all IPC omponents during recent evaluation


```{r all_ipcs}

data_components_all <- data_nziza_descriptive_recent_ipc_cs[,c(156:167)]

data_brut <- data_nziza_descriptive_recent_ipc_cs[,c(1:155, 170)]

dat_1 <- data_brut
dat_1$score_comp <- data_components_all$score_c_1
dat_1$ipc_component <- rep("comp_1", nrow(dat_1))


dat_2 <- data_brut
dat_2$score_comp <- data_components_all$score_c_2
dat_2$ipc_component <- rep("comp_2", nrow(dat_2))


dat_3 <- data_brut
dat_3$score_comp <- data_components_all$score_c_3
dat_3$ipc_component <- rep("comp_3", nrow(dat_3))


dat_4 <- data_brut
dat_4$score_comp <- data_components_all$score_c_4
dat_4$ipc_component <- rep("comp_4", nrow(dat_4))


dat_5 <- data_brut
dat_5$score_comp <- data_components_all$score_c_5
dat_5$ipc_component <- rep("comp_5", nrow(dat_5))


dat_6 <- data_brut
dat_6$score_comp <- data_components_all$score_c_6
dat_6$ipc_component <- rep("comp_6", nrow(dat_6))


dat_7 <- data_brut
dat_7$score_comp <- data_components_all$score_c_7
dat_7$ipc_component <- rep("comp_7", nrow(dat_7))


dat_8 <- data_brut
dat_8$score_comp <- data_components_all$score_c_8
dat_8$ipc_component <- rep("comp_8", nrow(dat_8))


dat_9 <- data_brut
dat_9$score_comp <- data_components_all$score_c_9
dat_9$ipc_component <- rep("comp_9", nrow(dat_9))


dat_10 <- data_brut
dat_10$score_comp <- data_components_all$score_c_10
dat_10$ipc_component <- rep("comp_10", nrow(dat_10))


dat_11 <- data_brut
dat_11$score_comp <- data_components_all$score_c_11
dat_11$ipc_component <- rep("comp_11", nrow(dat_11))


dat_12 <- data_brut
dat_12$score_comp <- data_components_all$score_c_12
dat_12$ipc_component <- rep("comp_12", nrow(dat_12))


dat_all <- rbind(dat_1, dat_2, dat_3, dat_4, dat_5, dat_6,
                 dat_7, dat_8, dat_9, dat_10, dat_11, dat_12)

# dat_all$score_comp <- dmm::unfactor(dat_all$score_comp)

ggplot(dat_all,aes(x=date_enquete,y=score_comp, fill=ipc_component))+
  geom_bar(stat="identity")+
  xlab("Date of evaluation") +
  ylab("Score") +
  ggtitle("Scores for all IPC omponents during recent evaluation")


```



## IPC scores for all IPC components per Health Zone during recent evaluation


```{r ipcs_hz}

ggplot(dat_all,aes(x=date_enquete,y=score_comp, fill=ipc_component))+
  geom_bar(stat="identity")+
  facet_wrap(zone_de_sante ~.)+
  xlab("Date of evaluation") +
  ylab("Score") +
  ggtitle("IPC scores for all IPC components per Health Zone during recent evaluation")

```



## IPC scores for all IPC components by presence of partners during recent evaluation


```{r ipc_part}

ggplot(dat_all,aes(x=date_enquete,y=score_comp, fill=ipc_component))+
  geom_bar(stat="identity")+
  facet_wrap(appuyee_par_partenaire ~.)+
  xlab("Date of evaluation") +
  ylab("Score") +
  ggtitle("IPC scores for all IPC components  by presence of partners during recent evaluation")


```


# Descriptive analysis for partners in IPC


## Supported Health Zones and their IPC score



```{r descr_hz}

partners_data <- dplyr::filter(data_nziza_descriptive_recent, !is.na(data_nziza_descriptive_recent$nbre_partenaires))

# transform in numeric 

partners_data_only <- partners_data[,27:80]


for(i in 1:ncol(partners_data_only)){
  
  partners_data_only[,i] <- as.numeric(partners_data_only[,i])
  
}


partners_data[,27:80] <- partners_data_only


partners_all_in_hz <- partners_data %>% 
  group_by(zone_de_sante) %>% 
  count(partenaire_adeco,
        partenaire_alima,
        partenaire_cac,
        partenaire_care_international,
        partenaire_caritas,
        partenaire_caritas_butembo,
        partenaire_cbca,
        partenaire_cdc_africa,
        partenaire_ceprossan,
        partenaire_cordaid,
        partenaire_famamundi,
        partenaire_fardc,
        partenaire_ima,
        partenaire_imc,
        partenaire_imt_anvers,
        partenaire_inrb,
        partenaire_institut_pasteur_de_dakar_ipd, 
        partenaire_irc,
        partenaire_irc_montpelier,
        partenaire_jica,
        partenaire_medair,
        partenaire_mercy_corps,
        partenaire_ministere_de_la_sante,
        partenaire_monusco,
        partenaire_mouvement_croix_rouge,
        partenaire_msf,
        partenaire_nih,
        partenaire_oim,
        partenaire_oms,
        partenaire_ong_musaka,
        partenaire_oxfam,
        partenaire_pam,
        partenaire_pap_rdc,
        partenaire_pdss,
        partenaire_pnc,
        partenaire_ppsp,
        partenaire_predict_metabiota,
        partenaire_premier_secours,
        partenaire_protection_civile,
        partenaire_racoj,
        partenaire_rb_fondation,
        partenaire_save_the_children,
        partenaire_sfcg,
        partenaire_sos_eau_et_foret,
        partenaire_tearfund,
        partenaire_ucla_drc,
        partenaire_umir,
        partenaire_unesco,
        partenaire_unfpa,
        partenaire_unhas,
        partenaire_unicef,
        partenaire_usamrid,
        partenaire_world_vision)


col_names <- names(partners_all_in_hz)
col_names <- gsub("partenaire_","",col_names)

names(partners_all_in_hz) <- col_names

partners_all_in_hz_groupe <- partners_all_in_hz

summary_partners_all_in_hz_groupe <- partners_all_in_hz_groupe %>% group_by(zone_de_sante) %>% tally()

names(summary_partners_all_in_hz_groupe) <- c("zone_de_sante", "nbr_partners")

# summary_partners_all_in_hz_groupe %>%
#   kable() %>%
#   kable_styling("striped") %>%
#   kableExtra::add_header_above( c("Table of Health Zone and number of IPC partners" = 2) ) %>%
#   kableExtra::row_spec(12)



# HZ Score nbre_partenaires

# score in HZ
stats_overall_partners_data <- ggpubr::desc_statby(partners_data, measure.var = "score_pci_100",
                                                   grps = "zone_de_sante")
stats_overall_partners_data_c <- stats_overall_partners_data[, c("zone_de_sante", "min", "max", "median", "mean", "iqr")]

stats_overall_partners_data_c[,2:6] <- round(stats_overall_partners_data_c[,2:6])


# Number of partners in these HZ
zone_de_sante_n_part <- partners_data %>% group_by(zone_de_sante) %>% tally()


stats_overall_partners_data_c_ns <- cbind(stats_overall_partners_data_c[,2:6], summary_partners_all_in_hz_groupe[,2])



stats_overall_partners_data_c_ns <- cbind(summary_partners_all_in_hz_groupe[,1], stats_overall_partners_data_c_ns) # adding name of HZ

names(stats_overall_partners_data_c_ns) <- c("zone_de_sante", "min", "max", "median", "mean", "iqr",  "nbr_partners")


stats_overall_partners_data_c_ns %>%
  kable() %>%
  kable_styling("striped") %>%
  kableExtra::add_header_above(c( "IPC Score and IPC partners and Health Facilities supported" = 7) ) %>%
  kableExtra::row_spec(12)



```



## IPC partners in Health Zones



```{r partner}

partners_all_in_hz %>%
  kable() %>%
  kable_styling("striped") %>%
  kableExtra::add_header_above( c("Table of Health Zone and IPC partners" = 55) ) %>%
  kableExtra::row_spec(74)

```



# Descriptive analysis of the difference between recent and previous IPC evaluations


Difference = Recent - Previous



```{r recent_previous}


rec_prev_dat <- recent_previous_fosa_evaluations(data_hz = data_scorecard_clean)


hz_evaluations <- rec_prev_dat %>% mutate(diffVal = scores_recent - scores_previous)


hz_evaluations_1 <- hz_evaluations

hz_evaluations_1$partner_status <- rep(NA, nrow(hz_evaluations_1))

for(i in 1:nrow(hz_evaluations_1)){
  
  if(hz_evaluations_1$partenaires_recent[i]==0){
    
    hz_evaluations_1$partner_status[i] <- "no"
    
  }else{
    
    hz_evaluations_1$partner_status[i] <- "yes"
    
  }
  
}

hz_evaluations <- hz_evaluations_1


```






## IPC score difference between recent and previous evaluations

```{r diff_recent_previous}

# Overall IPC score difference -------------

ggplot(data=hz_evaluations, aes(hz_evaluations$diffVal)) + 
  geom_histogram()+
  ggtitle("IPC score difference between recent and previous evaluations")+
  xlab("IPC score difference") +
  ylab("Frequency")

```



## IPC score difference with presence of partners



```{r diff_partn}

ggplot(data=hz_evaluations, aes(hz_evaluations$diffVal, fill=partner_status)) + 
  geom_histogram()+
  ggtitle("IPC score difference between recent and previous evaluation and partners")+
  xlab("IPC score difference") +
  ylab("Frequency")


```



## IPC score difference with presence of parteners by Health Zones


```{r diff_part_hz}


ggplot(data=hz_evaluations, aes(hz_evaluations$diffVal, fill=partner_status)) + 
  geom_histogram()+
  facet_wrap(nom_zone_sante_hz ~.)+
  ggtitle("IPC score difference between recent and previous evaluations in Health Zones and partners")+
  xlab("IPC score difference") +
  ylab("Frequency")



```




## IPC score difference by categories of Health Facilities



```{r diff_cat}

ggplot(data=hz_evaluations, aes(hz_evaluations$diffVal, fill = categorie_fosa)) + 
  geom_histogram()+
  ggtitle("IPC score difference between recent and previous evaluation")+
  xlab("IPC score difference") +
  ylab("Frequency")



## XXXXXXXXXXXXXXXXXXXX -------

```


# Descriptive analysis of the difference between recent and first IPC evaluations


Difference = Recent - First



```{r recent_first}


rec_prev_dat <- recent_first_fosa_evaluations(data_hz = data_scorecard_clean)


hz_evaluations <- rec_prev_dat %>% mutate(diffVal = scores_recent - scores_first)


hz_evaluations_1 <- hz_evaluations

hz_evaluations_1$partner_status <- rep(NA, nrow(hz_evaluations_1))

for(i in 1:nrow(hz_evaluations_1)){
  
  if(hz_evaluations_1$partenaires_recent[i]==0){
    
    hz_evaluations_1$partner_status[i] <- "no"
    
  }else{
    
    hz_evaluations_1$partner_status[i] <- "yes"
    
  }
  
}

hz_evaluations <- hz_evaluations_1


```






## IPC score difference between recent and first evaluations


```{r diff_recent_first}

# Overall IPC score difference -------------

ggplot(data=hz_evaluations, aes(hz_evaluations$diffVal)) + 
  geom_histogram()+
  ggtitle("IPC score difference between recent and first evaluations")+
  xlab("IPC score difference") +
  ylab("Frequency")

```



## IPC score difference with presence of parteners



```{r diff_partn_f}

ggplot(data=hz_evaluations, aes(hz_evaluations$diffVal, fill=partner_status)) + 
  geom_histogram()+
  ggtitle("IPC score difference between recent and first evaluation and partners")+
  xlab("IPC score difference") +
  ylab("Frequency")


```



## IPC score difference with presence of parteners by Health Zones


```{r diff_part_hz_f}


ggplot(data=hz_evaluations, aes(hz_evaluations$diffVal, fill=partner_status)) + 
  geom_histogram()+
  facet_wrap(nom_zone_sante_hz ~.)+
  ggtitle("IPC score difference between recent and first evaluations in Health Zones and partners")+
  xlab("IPC score difference") +
  ylab("Frequency")



```




## IPC score difference by categories of Health Facilities



```{r diff_cat_f}

ggplot(data=hz_evaluations, aes(hz_evaluations$diffVal, fill = categorie_fosa)) + 
  geom_histogram()+
  ggtitle("IPC score difference between recent and first evaluation")+
  xlab("IPC score difference") +
  ylab("Frequency")


```






